---
title: "Post Model-Fitting Diagnostics"
date: "2026-01-09"
author: "David Wolfson"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
    theme: sandstone 

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Objective: This document is to evaluate the output of 1,000 model runs of simulated true parameters and fitted model estimates.

Model context:
The model being run is a stage-based Leftkovich formulation of a Burnham joint live-encounter model with 3 stages. Only 2 stages can be recognized at
the time of banding (juvenile and adult), and we are inferring a 3rd stage (subadult) using a latent parameter (pi), which is the proportion of marked
adults that are actually subadults at time of banding. Pi has annual variation over the 20 year dataset.

Diagnostic results include:
- Rhat values  
- Bias estimates (truth minus estimated parameter values) 
- Precision estimates (coverage of credible intervals summarized across sampling runs vs truth)
- RMSE (mainly accuracy but also considers precision)

Changes from previous diagnostic documents:
- pi varies annually in the current model version
- calculated brownie's f (harvest index) directly during each model run
 

```{r, message=F, warning=F}
library(tidyverse)
library(here)
library(gridExtra)
library(patchwork)
```

### First, compile a dataset of all the true parameters simulated for each model run

All the file paths to saved RDS files.

```{r}
truth_files<-list.files(path=here("output/MSI_output/Jan_2026_lefkovitch/true_params"), pattern="*.rds", full.names = T)
```

Function to pull in all the saved RDS files and grab all the relevant information. 

```{r}
process_simulation_outputs <- function(paths) {
  # Extract the first integer sequence in the filename
  extract_run_number <- function(path) {
    as.integer(str_extract(basename(path), "\\d+"))
  }
  
  # Read all sims into a list of lists, keeping track of each simulation run number
  sims <- map(paths, ~{
    list(run = extract_run_number(.x),
         data = readRDS(.x))
  })
  
  # There are two kinds of simulated parameters: vectors, such as s, p, etc, which have 3 values, one for each age class,
  # and also the year-specific mu values, such as p.mu, s.mu, which are 3x20 matrices
  
  # Process 1D named vectors
  process_vector <- function(param) {
    map_dfr(sims, ~ tibble(
      run   = .x$run,                       # keep track of simulation run number
      index = seq_along(.x$data[[param]]),  # index represents age class (1=HY, 2=SY, 3=ASY)
      value = .x$data[[param]],             # actual truth value
      param = param                         # parameter name
    ))
  }
  
  # Process 2D matrices (make unique col names first)
  process_matrix <- function(param) {
    map_dfr(sims, function(sim) {
      mat <- sim$data[[param]]   # extract the matrix for the parameter of interest
      df  <- as.data.frame(mat)  # convert to a data frame
      
      colnames(df) <- paste0("V", seq_len(ncol(df))) # give them guaranteed unique names V1…Vn
      df %>%
        mutate(row = row_number()) %>%  # add row number as a column
        pivot_longer(                   # convert from wide to long format
          cols        = starts_with("V"),
          names_to    = "col",           # year number as col (i.e. column number from matrix)
          values_to   = "value"          # actual truth value
        ) %>%
        mutate(
          col   = as.integer(str_remove(col, "V")),     # add on year number as col variable
          run   = sim$run,                              # add simulation run number
          param = param                                 # add parameter name
        ) %>%
        select(run, row, col, value, param)
    })
  }
  
  # specify which parameters are vectors vs. matrices
  vector_params <- c("s.mu","s.sd","F.mu","F.sd","r.mu","r.sd","p.mu","p.sd","pi", "pi.mu", "pi.sd", "g", "g.mu", "g.sd") # pi.mu, pi.sd, g.mu, g.sd don't vary by stage (same value repeated) and have a static value across years, but still ok to grab
  matrix_params <- c("s","F","r","p") # f not present because it's derived, not simulated
  
  # build the result list
  results <- list()
  for (p in vector_params) results[[p]] <- process_vector(p)
  for (p in matrix_params) results[[p]] <- process_matrix(p)
  
  results
}
```



Extract info from rds file and combine everything into a list separated by parameter names.
```{r}
true_parameters<-process_simulation_outputs(truth_files)
```

### Now extract estimated model parameters from a csv saved from each model run.

Grab file paths.
```{r}
model_summary_files<-list.files(path=here("output/MSI_output/Jan_2026_lefkovitch/model_results"),
                                pattern="model_summary_simulation_run_.*\\.csv", 
                                full.names = T)
```

Function to import data separated by run number and parameter names.
```{r}
read_func<-function(paths){
  # Helper: extract the first integer sequence in the filename
  run_number<-as.integer(str_extract(basename(paths), "\\d+"))
  
  # read in csv file
  out<-read_csv(paths)
  
  
  # add in parameter names
  out<-data.frame(run=rep(run_number, nrow(out)), out)
}
```


Combine estimated model parameter values. (as_tibble output suppressed)
```{r, include=F}
model_df<-map_dfr(model_summary_files, read_func) %>% 
  as_tibble()
```

---

### Check R-hat values
This shows the relative convergence of all parameters.
```{r}
df<-model_df %>% 
  mutate(param_category=str_extract(param, "^[^\\[]+"),
         index=as.integer(str_extract(param, "(?<=\\[)\\d+(?=\\])")))

ggplot(df, aes(param_category, Rhat))+
  geom_boxplot(outliers = F)+
  geom_hline(yintercept=1.05, color="red")+
  labs(x="parameters")+
  theme_minimal()
#ggsave(plot=rhats, here("figures/rhats.jpg"), dpi=600)


# Table version

# df %>% 
#   group_by(param_category) %>% 
#   summarise(
#       mean_rhat = mean(Rhat, na.rm = TRUE),
#     sd_rhat = sd(Rhat, na.rm = TRUE),
#     quant_2.5=quantile(Rhat, prob=0.025),
#     quant_97.5=quantile(Rhat, prob=0.975)
#   )
```


---

## Coverage
Defined as the probability that the 95% CI from the fitted model parameter contains the true simulated value.

### Data-wrangling for coverage estimation

First get all the true parameters out of the big combined list and into separate dimension dataframes to compare.
True model parameters
```{r}
scalar_params <- c("pi.mu", "pi.sd", "g.mu", "g.sd") 
vector_params <- c("s.mu","s.sd","F.mu","F.sd","r.mu","r.sd","p.mu","p.sd", "pi", "g")
matrix_params <- c("s","F","r","p")
```


Get a logical index of names to filter from.
```{r}
vec_names<-names(true_parameters)%in%vector_params
matrix_names<-names(true_parameters)%in%matrix_params
scalar_names<-names(true_parameters)%in%scalar_params
```

Split up big single list of true parameters into categories of scalars, vectors, and matrices.
```{r}
true_vec_list <- true_parameters[vec_names]
truth_matrix_list <- true_parameters[matrix_names]
truth_scalars<-true_parameters[scalar_names]
```


Hypers for pi and g don't have 3 separate indices for stage, and don't vary by year, just a single value per run
```{r}
truth_scalars_df<-bind_rows(truth_scalars, .id="param")%>%
  rename(truth = value)
```

There are 2 different sets of dimensions to the vector parameters. 
Hyperpriors for the parameters that vary by stage are 3x1 vectors, with the 'index' variable keeping track of stage.
The parameters for 'g' and 'pi' are 20x1 vectors, with the 'index' variable keeping track of year.

 
It probably makes sense to keep index only as a variable for the different stages, and only use year for time-varying parameters.

First I'll parse out the hyperpriors that only vary by stage.
```{r}
hyper_vec_list<-c("s.mu","s.sd","F.mu","F.sd","r.mu","r.sd","p.mu","p.sd")

truth_hyper_vec_df<-bind_rows(true_vec_list, .id="param") %>% 
  filter(param%in%hyper_vec_list) %>% 
  rename(truth = value) %>% 
  mutate(year=NA) # so that everything can have a year column eventually
```

Now only by year. I'll rename index to year (because that is what it's representing), and give it another variable called index that is NA for record-keeping.
```{r}
yearly_vec_list<-c("pi", "g")

truth_year_vec_df<-bind_rows(true_vec_list, .id="param") %>% 
  filter(param%in%yearly_vec_list) %>% 
  rename(truth = value) %>% 
  rename(year=index) %>% 
  mutate(index=NA)
```

Combine the two kinds of vector parameters
```{r}
truth_vec_df<-bind_rows(truth_hyper_vec_df, truth_year_vec_df)
```


Parameters with 3x20 matrix for age variation plus temporal variation (20 years)
```{r}
truth_matrix_df<-bind_rows(truth_matrix_list, .id="param") %>% 
  rename(index=row, year=col, truth=value)
```

---

#### Now join together the combinations of true and estimated parameters

Model estimates for vector parameters

I'll follow the same process as before for breaking up the 3x1 vectors and 20x1 vectors and accounting for index vs year.

```{r}
est_hyper_vec_df <- model_df %>%
  mutate(
    index = as.integer(str_extract(param, "(?<=\\[)\\d+")),  # extract the number inside the brackets
    param = str_extract(param, "^[^\\[]+")   # grab the text before the left bracket
  ) %>%
  filter(param %in%hyper_vec_list) %>%
  select(run, param, index, mean, sd, X2.5., X50., X97.5., Rhat, n.eff) %>% 
  mutate(year=NA)

est_yearly_vec_df <- model_df %>% 
   mutate(
    index = as.integer(str_extract(param, "(?<=\\[)\\d+")),  # extract the number inside the brackets
    param = str_extract(param, "^[^\\[]+")   # grab the text before the left bracket
  ) %>%
  filter(param %in%yearly_vec_list) %>%
  select(run, param, index, mean, sd, X2.5., X50., X97.5., Rhat, n.eff) %>% 
  rename(year=index) %>% 
  mutate(index=NA)

est_vec_df<-bind_rows(est_hyper_vec_df, est_yearly_vec_df)
```

Scalar pi parameters. They don't vary by year or stage, single value per run.
Probably makes more sense to code index as NA, since the scalars are for all stages instead of a single one, than give it a numeric value.
```{r}
# estimated
scalar_est<-model_df %>% 
  filter(param%in%scalar_params)

# combine with true
comp_scalar<-scalar_est %>% 
  left_join(truth_scalars_df %>% 
              select(run, param, truth),
            by=c("run", "param")) %>% 
  mutate(index=NA) %>% 
  select(run, param, index, mean, sd, X2.5., X50., X97.5., Rhat, n.eff, truth) %>% 
  mutate(year=NA) %>% 
  distinct()
```


Scalar (pi) and vector parameters combined

```{r}
comp_vec <- left_join(est_vec_df, truth_vec_df, by = c("run", "index", "param", "year")) %>% 
  bind_rows(., comp_scalar) %>% 
  arrange(run, param, index, year) %>% 
  select(run, param, index, year, truth, mean, sd, X2.5., X50., X97.5., Rhat, n.eff)
```


Pull off indices from jags parameter names
```{r}
est_matrix_df<-model_df %>% 
  extract(
    col   = param,
    into  = c("param", "index", "year"),
    regex = "^([^\\[]+)\\[(\\d+),(\\d+)\\]$",  # first number goes to age index (1-3), second to year value (1-20)
    convert = TRUE
  ) %>% 
  filter(param%in%matrix_params) %>% 
  select(run, param, index, year, mean, sd, X2.5., X50., X97.5., Rhat, n.eff)
```


Add truth values to matrix parameter dataframe
```{r}
comp_mat <- left_join(est_matrix_df, truth_matrix_df) %>% 
  select(run, param, index, year, truth, mean, sd, X2.5., X50., X97.5., Rhat, n.eff)
```


Combine full set of truth and estimated to compare
```{r}
full_comp<-rbind.data.frame(comp_vec, comp_mat)
```


---

#### Now actually calculate and visualize coverage for hyperparameters, grouped by age index

Add approximate (Wald) CIs via standard error. Since sample size is large it's ok to use normal approximation.
```{r}
hypers<-c('F.mu', 'F.sd', 'p.mu', 'p.sd', 'r.mu', 'r.sd', 's.mu', 's.sd', 'pi')

param_coverage_approx <- full_comp %>% 
  filter(param%in%hypers) %>% 
  mutate(covered = (truth >= X2.5.) & (truth <= X97.5.)) %>% 
  group_by(param) %>% 
  summarise(
    index = index,
    coverage = mean(covered),
    n        = n()
  ) %>% 
  mutate(
    se       = sqrt(coverage * (1 - coverage) / n),
    ci_lower = coverage - 1.96 * se,
    ci_upper = coverage + 1.96 * se
  )

# So as to not jam too much together, I'll make a separate plot for each set of parameters

# give index numbers for age more descriptive labels
age_labels<-c(
  "1"="Juvenile",
  "2"="Subadult",
  "3"="Adult"
)

# true survival
survival_coverage<-param_coverage_approx %>% 
  filter(param%in%c('s.mu', 's.sd')) %>% 
ggplot(., aes(param, coverage)) +
  facet_wrap(~index, labeller=as_labeller(age_labels)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  geom_hline(yintercept = 0.95, color = "red") +
  theme_minimal()+
  ggtitle('Coverage of true survival (s) hyperparameters')+
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.text=element_text(size=16),
        strip.text=element_text(size=16),
        axis.title.y=element_text(size=16))+
  labs(x="", y="Coverage\n")

#ggsave(plot=survival_coverage, here("figures/survival_coverage.jpg"), dpi=600)


# dead encounter
recovery_coverage<-param_coverage_approx %>% 
  filter(param%in%c('r.mu', 'r.sd')) %>% 
ggplot(., aes(param, coverage)) +
  facet_wrap(~index, labeller=as_labeller(age_labels)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  geom_hline(yintercept = 0.95, color = "red") +
  theme_minimal()+
  ggtitle('Coverage of dead encounter (r) hyperparameters')+
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.text=element_text(size=16),
        strip.text=element_text(size=16),
        axis.title.y=element_text(size=16))+
  labs(x="", y="Coverage\n")

#ggsave(plot=recovery_coverage, here("figures/recovery_coverage.jpg"), dpi=600)



# # live encounter
# param_coverage_approx %>% 
#   filter(param%in%c('p.mu', 'p.sd')) %>% 
#   filter(index%in%c('2', '3')) %>% 
# ggplot(., aes(param, coverage)) +
#   facet_wrap(~index, labeller=as_labeller(age_labels)) +
#   geom_point() +
#   geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
#   geom_hline(yintercept = 0.95, color = "red") +
#   theme_minimal()+
#   ggtitle('Coverage of live encounter (p) hyperparameters')
# 
# 
# 
# # site fidelity
# param_coverage_approx %>% 
#   filter(param%in%c('F.mu', 'F.sd')) %>% 
# ggplot(., aes(param, coverage)) +
#   facet_wrap(~index, labeller=as_labeller(age_labels)) +
#   geom_point() +
#   geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
#   geom_hline(yintercept = 0.95, color = "red") +
#   theme_minimal()+
#   ggtitle('Coverage of site fidelity (F) hyperparameters')

# finite mixture models
pi_coverage<-param_coverage_approx %>% 
  filter(param=='pi') %>% 
ggplot(., aes(param, coverage)) +
  facet_wrap(~index, labeller=as_labeller(age_labels)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  geom_hline(yintercept = 0.95, color = "red") +
  theme_minimal()+
  ggtitle('Coverage of pi parameters')

ggsave(plot=pi_coverage, here("figures/pi_coverage.jpg"), dpi=600)

```
---

## Plot bias (truth minus estimates)

I'll first show my methods for a single parameter and then wrap everything in a function and visualize multiple parameters at the same time.




```{r, eval=F}
# Pull survival from truth
s.mu_truth<-true_parameters$s.mu %>%
  select(run, index, value) %>%
  rename(truth=value)

# Pull survival from model estimates
model_df_clean <- model_df %>%
  mutate(
    index = as.integer(str_extract(param, "(?<=\\[)\\d+")),  # extract index before modifying param
    param = str_extract(param, "^[^\\[]+")                   # extract base parameter name
  )

# Combine and calculate bias
s.mu_est<-model_df_clean %>%
  filter(param=="s.mu") %>%
  select(run, index, mean, sd)

s.mu_comp<-left_join(s.mu_est, s.mu_truth)

s.mu_comp<-s.mu_comp %>%
  mutate(bias=plogis(mean)-plogis(truth),
         rmse=sqrt(sd^2)+bias^2)

# Example for calculating 95% CI of bias across runs and visualizing 
ci_emp<-s.mu_comp %>%
  group_by(index) %>%
  summarise(
    lower = quantile(bias, 0.025, na.rm = TRUE),
    upper = quantile(bias, 0.975, na.rm = TRUE)
  )

# bias
ggplot(s.mu_comp, aes(x = factor(index), y = bias)) +
  geom_violin(trim = FALSE, fill = "lightgreen") +
  geom_linerange(
    data = ci_emp,
    inherit.aes = F,
    aes(x = factor(index), ymin = lower, ymax = upper),
    color = "darkgreen", linewidth = 1
  ) +
  stat_summary(fun = median, geom = "point", color = "black", size = 3.5) +
  geom_hline(yintercept = 0, color="red")+
  labs(title = "Bias violin plots with 2.5% and 97.5% quantiles",
       x = "Age class (index)", y = "Bias") +
  theme_minimal()+
  scale_x_discrete(labels=c("HY", "SY", "ASY"))

# now for rmse
ci_emp_rmse<-s.mu_comp %>%
  group_by(index) %>%
  summarise(
    lower = quantile(rmse, 0.025, na.rm = TRUE),
    upper = quantile(rmse, 0.975, na.rm = TRUE)
  )



ggplot(s.mu_comp, aes(x = factor(index), y = rmse)) +
  geom_violin(trim = FALSE, fill = "lightgreen") +
  geom_linerange(
    data = ci_emp_rmse,
    inherit.aes = F,
    aes(x = factor(index), ymin = lower, ymax = upper),
    color = "darkgreen", linewidth = 1
  ) +
  stat_summary(fun = median, geom = "point", color = "black", size = 3.5) +
  geom_hline(yintercept = 0, color="red")+
  labs(title = "RMSE violin plots with 2.5% and 97.5% quantiles",
       x = "Age class (index)", y = "RMSE") +
  theme_minimal()+
  scale_x_discrete(labels=c("HY", "SY", "ASY"))


```


#### Now everything together. First as a function to recreate all the above steps, and then with an additional wrapper to simultaneously do the same thing for multiple parameters at the same time.

One parameter function
```{r}
plot_bias_violin_one <- function(param_name, model_df, true_parameters) {
  truth_df <- true_parameters[[param_name]] %>%
    select(run, index, value) %>%
    rename(truth = value)
  
  est_df <- model_df %>%
    mutate(
      index = as.integer(str_extract(param, "(?<=\\[)\\d+")),
      param = str_extract(param, "^[^\\[]+")
    ) %>%
    filter(param == param_name) %>%
    select(run, index, mean, sd, X2.5., X97.5.)
  
  comp_df <- left_join(est_df, truth_df, by = c("run", "index")) %>%
    mutate(bias = plogis(mean) - plogis(truth),
         rmse=sqrt(sd^2+bias^2))
  
  ci_emp <- comp_df %>%
    group_by(index) %>%
    summarise(
      lower = quantile(bias, 0.025, na.rm = TRUE),
      upper = quantile(bias, 0.975, na.rm = TRUE),
      .groups = "drop"
    )
  
  ggplot(comp_df, aes(x = factor(index), y = bias)) +
    geom_violin(trim = T, fill = "lightgreen", color=NA) +
    geom_linerange(
      data        = ci_emp,
      inherit.aes = FALSE,
      aes(x = factor(index), ymin = lower, ymax = upper),
      linewidth   = 1.2,
      color       = "darkgreen"
    ) +
    stat_summary(fun = median, geom = "point", color = "black", size = 3) +
    geom_hline(yintercept = 0, color = "red") +
    labs(
      title = paste0("Bias for ‘", param_name, "’"),
      x     = "Age class",
      y     = "Bias"
    ) +
    theme_minimal()+
    theme(plot.title=element_text(hjust=0.5),
          axis.text.x=element_text(size=16))+
      scale_x_discrete(labels=c("HY", "SY", "ASY"))  # this will make an error when HY p is taken out, but otherwise it's more convenient
}

plot_rmse_violin_one<- function(param_name, model_df, true_parameters) {
  truth_df <- true_parameters[[param_name]] %>%
    select(run, index, value) %>%
    rename(truth = value)
  
  est_df <- model_df %>%
    mutate(
      index = as.integer(str_extract(param, "(?<=\\[)\\d+")),
      param = str_extract(param, "^[^\\[]+")
    ) %>%
    filter(param == param_name) %>%
    select(run, index, mean, sd)
  
  comp_df <- left_join(est_df, truth_df, by = c("run", "index")) %>%
    mutate(bias = plogis(mean) - plogis(truth),
         rmse=sqrt(sd^2)+bias^2)
  
  ci_emp_rmse <- comp_df %>%
    group_by(index) %>%
    summarise(
      lower = quantile(rmse, 0.025, na.rm = TRUE),
      upper = quantile(rmse, 0.975, na.rm = TRUE),
      .groups = "drop"
    )
  
  ggplot(comp_df, aes(x = factor(index), y = rmse)) +
  geom_violin(trim = T, fill = "lightgreen", color=NA) +
  geom_linerange(
    data = ci_emp_rmse,
    inherit.aes = F,
    aes(x = factor(index), ymin = lower, ymax = upper),
    color = "darkgreen", linewidth = 1.2
  ) +
  stat_summary(fun = median, geom = "point", color = "black", size = 3) +
    geom_hline(yintercept = 0, color = "red") +
    labs(
      title = paste0("RMSE for ‘", param_name, "’"),
      x     = "Age class",
      y     = "RMSE"
    ) +
    theme_minimal()+
    theme(plot.title=element_text(hjust=0.5),
                 axis.text.x=element_text(size=16))+
      scale_x_discrete(labels=c("HY", "SY", "ASY"))  # this will make an error when HY p is taken out, but otherwise it's more convenient
}

```

Wrapper for multiple parameters
```{r}
plot_bias_violin_grid <- function(param_names, model_df, true_parameters,
                                  ncol = 2,  # number of columns in the grid
                                  plot_width = 6, plot_height = 4) {
  # generate one plot per parameter
  plot_list <- lapply(param_names, function(p) {
    plot_bias_violin_one(p, model_df, true_parameters)
  })
  
  # arrange them in a grid
  grid_obj <- do.call(grid.arrange,
                      c(plot_list, ncol = ncol))
  
  # Return invisibly so that you can capture it if desired
  invisible(grid_obj)
}

plot_rmse_violin_grid <- function(param_names, model_df, true_parameters,
                                  ncol = 2,  # number of columns in the grid
                                  plot_width = 6, plot_height = 4) {
  # generate one plot per parameter
  plot_list <- lapply(param_names, function(p) {
    plot_rmse_violin_one(p, model_df, true_parameters)
  })
  
  # arrange them in a grid
  grid_obj <- do.call(grid.arrange,
                      c(plot_list, ncol = ncol))
  
  # Return invisibly so that you can capture it if desired
  invisible(grid_obj)
}
```


Vector parameters to compare
```{r}
mu_to_plot <- c("s.mu", "F.mu", "p.mu", "r.mu")
sd_to_plot <- c("s.sd", "F.sd", "p.sd", "r.sd")

reduced_mus<- c("s.mu", "r.mu")
```

Produce 2-column grid of the plots. The black dot in each figure is the median value and the dark green vertical line is the range of the 2.5% and 97.5% quantiles across simulation runs"
```{r}
# first exclude p1 since it isn't being estimated
df_filter<-model_df %>% 
  filter(!param%in%c('p.mu[1]', 'p.sd[1]'))


plot_bias_violin_grid(mu_to_plot, df_filter, true_parameters, ncol = 2)
plot_bias_violin_grid(sd_to_plot, df_filter, true_parameters, ncol = 2)

plot_rmse_violin_grid(mu_to_plot, df_filter, true_parameters, ncol = 2)
# I don't think it's appropriate/needed to plot rmse for sd

reduced_mu_bias_plot<-plot_bias_violin_grid(reduced_mus, df_filter, true_parameters, ncol = 2)
ggsave(plot=reduced_mu_bias_plot, here("figures/reduced_mu_bias_plot.jpg"), dpi=600)

reduced_mu_rmse_plot<-plot_rmse_violin_grid(reduced_mus, df_filter, true_parameters, ncol = 2)
ggsave(plot=reduced_mu_rmse_plot, here("figures/reduced_mu_rmse_plot.jpg"), dpi=600)

# grid version doesn't line up the y=0 line in the plots, which is weird.
# it's probably better to keep the bias and rmse for parameters together, so I'll call the single plot functions and then put plots together with cowplot or patchwork


s_bias<-plot_bias_violin_one('s.mu', df_filter, true_parameters)
s_rmse<-plot_rmse_violin_one('s.mu', df_filter, true_parameters)
surv_plots<-s_bias+s_rmse
ggsave(plot=surv_plots, here("figures/survival_bias_rmse.jpg"), dpi=600)

r_bias<-plot_bias_violin_one('r.mu', df_filter, true_parameters)
r_rmse<-plot_rmse_violin_one('r.mu', df_filter, true_parameters)
recovery_plots<-r_bias+r_rmse
ggsave(plot=recovery_plots, here("figures/recovery_bias_rmse.jpg"), dpi=600)


```


---

#### Now do something similar for the 3x20 matrix parameters

First a function to summarize bias and quantiles
```{r}
summarise_matrix_bias <- function(param_name, model_df, true_parameters) {
  df_mat <- model_df %>%
    # only keep rows like "s[1,1]" if param_name=="s"
    filter(str_detect(param, paste0("^", param_name, "\\["))) %>%
    extract(
      col     = param,
      into    = c("param_name", "row", "col"),
      regex   = "^(\\w+)\\[(\\d+),(\\d+)\\]$",
      convert = TRUE
    ) %>%
    select(run, row, col, mean)
  
  truth_df <- true_parameters[[param_name]] %>%
    rename(row = row, col = col, truth = value)
  
  comp <- df_mat %>%
    left_join(truth_df, by = c("run","row","col")) %>%
    mutate(bias = mean - truth)
  
  ci <- comp %>%
    group_by(row, col) %>%
    summarise(
      lower = quantile(bias, 0.025, na.rm = TRUE),
      upper = quantile(bias, 0.975, na.rm = TRUE),
      med   = median(bias,    na.rm = TRUE),
      .groups = "drop"
    )
  
  list(comp = comp, ci = ci)
}

summarise_matrix_rmse <- function(param_name, model_df, true_parameters) {
  df_mat <- model_df %>%
    # only keep rows like "s[1,1]" if param_name=="s"
    filter(str_detect(param, paste0("^", param_name, "\\["))) %>%
    extract(
      col     = param,
      into    = c("param_name", "row", "col"),
      regex   = "^(\\w+)\\[(\\d+),(\\d+)\\]$",
      convert = TRUE
    ) %>%
    select(run, row, col, mean, sd)
  
  truth_df <- true_parameters[[param_name]] %>%
    rename(row = row, col = col, truth = value)
  
  comp <- df_mat %>%
    left_join(truth_df, by = c("run","row","col")) %>%
    mutate(bias = mean - truth,
         rmse=sqrt(sd^2+bias^2))
  
  ci <- comp %>%
    group_by(row, col) %>%
    summarise(
      lower = quantile(rmse, 0.025, na.rm = TRUE),
      upper = quantile(rmse, 0.975, na.rm = TRUE),
      med   = median(rmse,    na.rm = TRUE),
      .groups = "drop"
    )
  
  list(comp = comp, ci = ci)
}
```



Violin plot style function
```{r}
plot_matrix_violins_bias <- function(param_name, model_df, true_parameters,
                                ncol = 3 # how many cols of facets
) {
  out  <- summarise_matrix_bias(param_name, model_df, true_parameters)
  comp <- out$comp
  ci   <- out$ci
  
  ggplot(comp, aes(x = factor(col), y = bias)) +
    geom_violin(trim = T, fill = "lightblue", color=NA) +
    stat_summary(fun = median, geom = "point", color = "black", size = 2) +
    geom_linerange(
      data        = ci,
      inherit.aes = FALSE,
      aes(x = factor(col), ymin = lower, ymax = upper),
      linewidth   = .4, color = "darkblue"
    ) +
    geom_hline(yintercept = 0, color = "red") +
    scale_x_discrete(breaks=c(1,5,10,15,20), 
                       labels=c('1', '5', '10', '15', '20'))+
    facet_wrap(~ row, ncol = ncol, labeller=as_labeller(age_labels)) +
    labs(
      title = paste0("Bias of ‘", param_name, "' by age class and year"),
      x     = "Year",
      y     = "Bias"  # could also add more detail with: "(mean – truth)"
    ) +
    theme_minimal()
}

plot_matrix_violins_rmse <- function(param_name, model_df, true_parameters,
                                     ncol=3) {
  out  <- summarise_matrix_rmse(param_name, model_df, true_parameters)
  comp <- out$comp
  ci   <- out$ci
  
  ggplot(comp, aes(x = factor(col), y = rmse)) +
    geom_violin(trim = T, fill = "lightblue", color=NA) +
    stat_summary(fun = median, geom = "point", color = "black", size = 2) +
    geom_linerange(
      data        = ci,
      inherit.aes = FALSE,
      aes(x = factor(col), ymin = lower, ymax = upper),
      linewidth   = 0.4, color = "darkblue"
    ) +
    geom_hline(yintercept = 0, color = "red") +
        scale_x_discrete(breaks=c(1,5,10,15,20), 
                       labels=c('1', '5', '10', '15', '20'))+
    facet_wrap(~ row, ncol = ncol, labeller=as_labeller(age_labels)) +
    labs(
      title = paste0("RMSE of ‘", param_name, "’ by age class and year"),
      x     = "Year",
      y     = "RMSE"  # to add the equation for RMSE use: bquote("RMSE" ~ sqrt(bias^2+variance^2))
    ) +
    theme_minimal()
}



```




Violin plot for the s (true survival) matrix
```{r}
s.yearly.bias<-plot_matrix_violins_bias("s", model_df, true_parameters)
ggsave(plot=s.yearly.bias, here("figures/yearly_survival_bias.jpg"),
       dpi=600)



s.yearly.rmse<-plot_matrix_violins_rmse("s", model_df, true_parameters)
ggsave(plot=s.yearly.rmse, here("figures/yearly_survival_rmse.jpg"),dpi=600)

yearly_survival_plot<-s.yearly.bias+s.yearly.rmse
ggsave(plot=yearly_survival_plot, here("figures/yearly_survival_bias_rmse.jpg"), dpi=600)

```





Violin plot for the r (dead encounter) matrix
```{r}
r.yearly.bias<-plot_matrix_violins_bias("r", model_df, true_parameters)
ggsave(plot=r.yearly.bias, here("figures/yearly_recovery_bias.jpg"), dpi=600)

r.yearly.rmse<-plot_matrix_violins_rmse("r", model_df, true_parameters)
ggsave(plot=r.yearly.rmse, here("figures/yearly_recovery_rmse.jpg"), dpi=600)
```



Violin plot for the F (site fidelity) matrix
```{r}
plot_matrix_violins_bias("F", model_df, true_parameters)
plot_matrix_violins_rmse("F", model_df, true_parameters)
```





Violin plot for the p (live encounter) matrix
```{r}
# params to drop
p1<-c("p[1,1]", "p[1,2]", "p[1,3]",
      "p[1,4]", "p[1,5]", "p[1,6]", 
      "p[1,7]", "p[1,8]", "p[1,9]",
      "p[1,10]", "p[1,11]", "p[1,12]",
      "p[1,13]", "p[1,14]", "p[1,15]",
      "p[1,16]",  "p[1,17]",  "p[1,18]",
      "p[1,19]", "p[1,20]")

# First a filtered dataset that drops HY p parameters
p_filt<-model_df %>% 
  filter(!param%in%p1)


plot_matrix_violins_bias("p", p_filt, true_parameters)
plot_matrix_violins_rmse("p", p_filt, true_parameters)
```



Calculate brownie's f (index of harvest rate)
```{r}
  
  # Pull survival from truth
s.mu_truth<-true_parameters$s.mu %>%
  select(run, index, value) %>%
  rename(truth=value) %>% 
  mutate(true_param="s")

# 'De-index'  model estimates
model_df_clean <- model_df %>%
  mutate(
    index = as.integer(str_extract(param, "(?<=\\[)\\d+")),  # extract index before modifying param
    param = str_extract(param, "^[^\\[]+")                   # extract base parameter name
  )

# pull r (dead encounter) from truth
r.mu_truth<-true_parameters$r.mu %>% 
  select(run, index, value) %>%
  rename(truth=value) %>% 
  mutate(true_param="r")

# calculate true brownie f from simulations
brownie.mu_truth<-rbind.data.frame(s.mu_truth, r.mu_truth)
brownie.mu_truth<-brownie.mu_truth %>% 
  pivot_wider(
    names_from = true_param,
    values_from=truth
  ) %>% 
  mutate(f_true=(1-s)*r)


# Calculate estimated brownie's f
brownie.mu_est<-model_df_clean %>%
  filter(param%in%c("s.mu", "r.mu")) %>%
  select(run, index, param, mean) %>% 
  pivot_wider(
    names_from=param,
    values_from = mean
  ) %>% 
  mutate(f_est=(1-s.mu)*r.mu)

brownie.mu_comp<-left_join(brownie.mu_est, brownie.mu_truth)

brownie.mu_comp<-brownie.mu_comp %>%
  mutate(bias_f=plogis(f_est)-plogis(f_true))

# # Example for calculating 95% CI across runs and visualizing 
ci_emp<-brownie.mu_comp %>%
  group_by(index) %>%
  summarise(
    lower = quantile(bias_f, 0.025, na.rm = TRUE),
    upper = quantile(bias_f, 0.975, na.rm = TRUE)
  )

brownie_f_plot<-ggplot(brownie.mu_comp, aes(x = factor(index), y = bias_f)) +
  geom_violin(trim = T, fill = "lightgreen", color=NA) +
  geom_linerange(
    data = ci_emp,
    inherit.aes = F,
    aes(x = factor(index), ymin = lower, ymax = upper),
    color = "darkgreen", linewidth = 1.2
  ) +
  stat_summary(fun = median, geom = "point", color = "black", size = 3) +
  geom_hline(yintercept = 0, color="red")+
  labs(title = "Brownie's f (harvest rate)",
       x = "Age class", y = "Bias") +
  theme_minimal()+
  scale_x_discrete(labels=c("HY", "SY", "ASY"))+
  theme(plot.title = element_text(hjust=0.5),
                 axis.text.x=element_text(size=14))

ggsave(plot=brownie_f_plot, here("figures/brownie_f_bias.jpg"), dpi=600)



# ggplot(brownie.mu_comp, aes(x = factor(index), y = bias_f)) +
#   geom_boxplot(fill = "lightgreen", outliers = F) +
#   geom_hline(yintercept = 0, color="red")+
#   labs(title = "Brownie's f (harvest rate)",
#        x = "Age class", y = "Bias") +
#   theme_minimal()+
#   scale_x_discrete(labels=c("HY", "SY", "ASY"))+
#   theme(plot.title = element_text(hjust=0.5))

```

Run pi separately
```{r}
  
  # Pull pi from truth
pi_truth<-true_parameters$pi %>%
  select(run, value) %>%
  rename(truth=value)


# 'De-index'  model estimates
pi_df_clean <- model_df %>%
  mutate(
    index = as.integer(str_extract(param, "(?<=\\[)\\d+")),  # extract index before modifying param
    param = str_extract(param, "^[^\\[]+")                   # extract base parameter name
  ) %>% 
    filter(param=="pi")



pi_comp<-left_join(pi_df_clean, pi_truth)

pi_comp<-pi_comp %>%
  mutate(bias=plogis(mean)-plogis(truth))

# # Example for calculating 95% CI across runs and visualizing 
ci_emp<-pi_comp %>%
  summarise(
    lower = quantile(bias, 0.025, na.rm = TRUE),
    upper = quantile(bias, 0.975, na.rm = TRUE)
  )

pi_bias_plot<-ggplot(pi_comp, aes(x='', y = bias)) +
  geom_violin(trim = T, fill = "lightgreen", color=NA) +
  geom_linerange(
    data = ci_emp,
    inherit.aes = F,
    aes(x='', ymin = lower, ymax = upper),
    color = "darkgreen", linewidth = 1.2
  ) +
  stat_summary(fun = median, geom = "point", color = "black", size = 3) +
  geom_hline(yintercept = 0, color="red")+
  labs(title = "Bias in pi parameter",
       y = "Bias") +
  theme_minimal()+
  theme(plot.title = element_text(hjust=0.5),
                 axis.text.x=element_text(size=14))

ggsave(plot=pi_bias_plot, here("figures/pi_bias.jpg"), dpi=600)





```




```{r}
sessionInfo()
```






